---
title: "instacart"
author: 
date: "09/03/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

The main goal of recommender systems is to provide suggestions to online users to make better decisions from many alternatives available over the Web. A better recommender system is directed more towards personalized recommendations by taking into consideration information about a product, such as specifications, purchase histor the users, comparison with other products, and so on, before making recommendations.

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(data.table)
library(tibble)
library(rsparse)
library(Matrix)
library(Metrics)
```


```{r}
#Import data

orders<-read.csv("orders.csv")
products<-read.csv("products.csv")
departments<-read.csv("departments.csv")
prior<-read.csv("order_products__prior.csv")
train<-read.csv("order_products__train.csv")
test<-read.csv("sample_submission.csv")



```
```{r include=FALSE}

#testing

head(orders)
#dim(orders)
orders %>% 
  filter(order_id==2)

orders %>% 
  filter(eval_set=="train")
```
```{r include=FALSE}

#testing
head(prior)
prior %>% 
  filter(order_id==2)

head(products)

```
```{r}
#merge prior & order 

prior_order<-orders %>% 
  filter(eval_set=="prior") %>% 
    left_join(prior,orders,by=c("order_id")) %>% 
    left_join(products,by=c("product_id")) %>% 
  filter(reordered==1)
#rm(orders,prior)

head(prior_order)

```

```{r inlcude=FALSE}

#testing
prior_order %>% 
  filter(user_id==107749) %>% 
  distinct(product_id)


products %>% 
  filter(product_id %in% c(43828,30833,28661,24507,21606,20032) )

products %>% 
  filter(product_id %in% c(8021	,			
27521		,		
34864	,			
5383	,			
49145	,			
3563	,			
16278	,			
8277	,			
42843	,		
19660) )

```


```{r include=FALSE}

#testing
prior_order%>% 
  filter(order_id==2196797)


prior_order %>% 
  filter(user_id==202279)
```

```{r}
#merge train and order

train_order<-orders %>% 
  filter(eval_set=="train") %>% 
    left_join(train,orders,by=c("order_id"))


```



```{r}
#merge test and order

test_order<-orders %>% 
  filter(eval_set=="test") %>% 
    left_join(test,orders,by=c("order_id"))


```
```{r}
dim(prior_order)
dim(train_order)
dim(test_order)
```


```{r}
## Total users 

user_count<-unique(orders$user_id)
length(user_count)

```


```{r}
#total products
product_count<-unique(products$product_id)
length(product_count)

```


```{r}
orders_count<-unique(orders$order_id)
length(orders_count)
```



```{r}
#total products in prior
product_prior_count<-unique(prior_order$product_id)
length(product_prior_count)

```

```{r}
#total products in train

product_train_count<-unique(train_order$product_id)
length(product_train_count)
```

```{r include=FALSE}
#top 50 products train

top_products_train<-train_order %>% 
 group_by(user_id,product_id) %>% 
 summarise(tot=n()) %>% 
  ungroup() %>% 
  group_by(product_id) %>% 
  summarise(count1=n()) %>% 
  ungroup() %>% 
arrange(desc(count1)) %>% 
top_n(50)

#top_products_train

```


```{r include=FALSE}

# train users for top 50  

train_users<-prior_order %>%
  filter(product_id %in% top_products_train$product_id ) %>% 
 group_by(user_id,product_id) %>% 
  summarise(tot=n()) %>% 
  ungroup() %>% 
  group_by(user_id) %>% 
  summarise(count2=n()) %>% 
 arrange(desc(count2)) 



#train_users

```
```{r}

# top 50  products prior
top_products_prior<-prior_order %>% 
 group_by(user_id,product_id) %>% 
 summarise(tot=n()) %>% 
  ungroup() %>% 
  group_by(product_id) %>% 
  summarise(count1=n()) %>% 
  ungroup() %>% 
arrange(desc(count1)) %>% 
top_n(50)


#top_products_prior

```




```{r}
#prior user for top 50

prior_users<-prior_order %>%
  filter(product_id %in% top_products_prior$product_id ) %>% 
 group_by(user_id,product_id) %>% 
   summarise(tot=n()) %>% 
  ungroup() %>% 
  group_by(user_id) %>% 
  summarise(count2=n()) %>% 
 arrange(desc(count2)) 



#prior_users
```

```{r include=FALSE}
#testing
# z<-prior_order %>% 
#   filter(product_id %in% top_products_prior$product_id & user_id==201142) %>% 
#   distinct(product_id,.keep_all = TRUE)
# z




```


```{r include=FALSE}

#transaction for top 50 products (train)

# train_order %>% 
# filter(user_id %in%train_users$user_id  & product_id %in% top_products_train$product_id)
```


```{r include=FALSE}

#transaction for top 50 products (prior)

# prior_order %>% 
# filter(user_id %in%prior_users$user_id  & product_id %in% top_products_prior$product_id)

```




```{r include=FALSE}
# top_products<-prior_order %>% 
#  group_by(user_id,product_id) %>% 
#  summarise(tot=n()) %>% 
#   ungroup() %>% 
#   group_by(product_id) %>% 
#   summarise(count1=n()) %>% 
#   ungroup() %>% 
# arrange(desc(count1))
#   
# top_products
```

```{r include=FALSE}
#test

# x<-prior_order %>% 
#   filter(product_id==24852) %>% 
#   distinct(user_id,.keep_all = TRUE)
# x
```


```{r}
#Number of distinct orders by user

p1<-orders %>% 
  group_by(user_id) %>% 
  summarise(count_order=n()) %>% 
  ungroup() 

# %>% group_by(count_order) %>% 
#   summarise(count_user=n())

ggplot(p1, aes(count_order)) + geom_histogram(binwidth = 10)+labs(title="Number of distinct orders by user",
        x ="Number of Orders", y = "Number of Users")+coord_cartesian(xlim = c(0, 80))


```

```{r}

#combine prior and train

all<-rbind(train_order,prior_order)
#head(all,100)
```

```{r}
#Number of distinct products/user

p2<-all %>% 
  group_by(user_id,product_id) %>% 
  summarise(count3=n()) %>% 
  select(user_id,product_id,count3) %>% 
  ungroup() %>% 
  group_by(user_id) %>% 
  summarise(count_product=n()) %>% 
  ungroup() 
# %>% group_by(count_product) %>% 
#   summarise(count_user=n())
  


ggplot(p2, aes(count_product)) + geom_histogram(binwidth = 10)+labs(title="Number of distinct products/user",
        x ="Number of Products", y = "Number of Users")+coord_cartesian(xlim = c(0, 250))


```

```{r}
#Number of distinct users/item



p3<-all %>% 
  group_by(product_id,user_id) %>% 
  summarise(count4=n()) %>% 
   select(user_id,product_id,count4) %>% 
  ungroup() %>%
  group_by(product_id) %>% 
  summarise(count_user=n()) %>% 
  ungroup() 
# %>% group_by(count_user) %>% 
#   summarise(count_product=n())


ggplot(p3, aes(count_user)) + geom_histogram(binwidth = 50)+labs(title="Number of distinct users/item",
        x ="Number of Users", y = "Number of Products")+coord_cartesian(xlim = c(0, 500))

 
```

```{r include=FALSE}


#Number of Orders by User/Product


# p4<-prior_order %>% 
#   group_by(user_id,product_id,order_id) %>% 
#   ungroup() %>% 
#   group_by(user_id,product_id) %>% 
#   summarise(count_order=n()) %>%
#   ungroup()%>% 
#   group_by(count_order) %>% 
# summarise(count_user_product=n()) %>% 
#   top_n(10)
# 
# 
# 
# p4$count_order<-as.factor(p4$count_order)
# 
# ggplot(p4, aes(count_order)) + geom_histogram(binwidth = 0.001)+labs(title="Number of Orders by User/Product",
#         x ="Number of Orders", y = "Number of Users/Products")+coord_cartesian(ylim = c(0, 8000000))
# 
# 

```
```{r include=FALSE}


#testing
# prior_order %>% 
#  filter( user_id==1 & product_id==196)
```

```{r}
rm(orders,products,departments,prior,train,test,test_order,train_order,user_count,product_count,orders_count)
rm( product_prior_count,product_train_count,top_products_train,train_users,top_products_prior,prior_users,p1,all,p2,p3)

#Extract columns for matrix
transactions<-prior_order[,c("user_id","product_id","order_id")]
set.seed(5)
#selecting 8K users for modeling
all_users<-unique(transactions$user_id)
randm_users<-sample(all_users,8000L)

set.seed(10)
#selecting 16K products
all_products<-unique(transactions$product_id)
rand_products<-sample(all_products,16000L)



#final data for matrix
interactions<-transactions %>% 
                     filter(user_id %in%randm_users & product_id %in%rand_products)

#find the total orders for each user per product
interactions_sample<-interactions %>% 
  group_by(user_id,product_id) %>% 
  summarise(orders=n())


# rm(transactions,interactions,all_users,prior_order,randm_users)
# 
# 
# dim(interactions)
# head(interactions_sample,100)
```


```{r }


#encoding users and products
user_enc <-  interactions_sample %>%
  distinct(user_id) %>%
 rowid_to_column()

names(user_enc)[names(user_enc) == "rowid"]<- "uid_enc"
  
product_enc<- unique(interactions_sample[,2]) %>% 
 rowid_to_column() 

names(product_enc)[names(product_enc) == "rowid"]<- "pid_enc"




#73958
#dim(product_enc)
#head(product_enc)
#7486
#h<-unique(product_enc$product_id)
#length(h)



# length(product_enc$product_id)
# dim(product_enc)
# head(product_enc,100)
# head(user_enc)
# 
# product_enc %>% 
#   filter(pid_enc==1)

```


```{r}
#sparse matrix
data<-interactions_sample %>%
  inner_join(user_enc, by = 'user_id') %>% 
  inner_join(product_enc, by = 'product_id')



X = sparseMatrix(i = data$uid_enc, j = data$pid_enc, x = data$orders, 
                 dimnames = list(user_enc$user_id, product_enc$product_id))







#rm(interactions_sample)
#rm(data)
# str(data)
# dim(data)
# head(data)
# dim(X)
# head(X)
# class(X)
# data %>% 
#   filter(uid_enc==1) %>% 
#   arrange(pid_enc)
# data %>% 
#   filter(pid_enc==1) %>% 
#   arrange(pid_enc)

```


```{r}
#test set

set.seed(100)
n_test <- 2000L
test_uid <- sample(nrow(user_enc), n_test)

X_train <-  X[-test_uid, ]
X_test <-  X[test_uid, ]

 
# dim(X_train)
# dim(X_test)
# class(test_uid)

```



```{r}
# Split our test set into "history" or "future"
temp = as(X_test, "TsparseMatrix")
temp = data.table(i = temp@i, j = temp@j, x = temp@x) 



temp <- temp %>%
  group_by(i) %>%                         # group by user
  mutate(ct = length(j),                  # number of products each user has
         history = 
           sample(c(TRUE, FALSE), ct, replace = TRUE, prob = c(.5, .5))) %>%
  select(-ct)


X_test_history <- temp %>% filter(history == TRUE)
X_test_future <- temp %>% filter(history == FALSE)




X_test_history <- sparseMatrix(i = X_test_history$i,
                               j = X_test_history$j,
                               x = X_test_history$x,
                               dims = dim(X_test),
                               dimnames = dimnames(X_test),
                               index1 = FALSE)



X_test_future <- sparseMatrix(i = X_test_future$i,
                              j = X_test_future$j,
                              x = X_test_future$x,
                              dims = dim(X_test),
                              dimnames = dimnames(X_test),
                              index1 = FALSE)

#rm(temp)
# rm(user_enc, product_enc, n_test, test_uid)
# dim(temp)
# dim(X_test_history)
# head(X_test_history)
# dim(X_test_future)
# head(X_test_future)
# colnames(X_test_history)
```




```{r}
#testing
# 
# products %>%
#   filter(product_id==16797)
# 
# 
#orders %>% 
 #  filter(user_id==88057)
# 
prior_order %>% 
filter(user_id==88057) %>% 
filter(product_id==21903)

```




```{r}
#  confidence functions and create matrices

lin_conf <- function(x, alpha) {
  x_confidence <- x
  stopifnot(inherits(x, "sparseMatrix"))
  x_confidence@x = 1 + alpha * x@x
  return(x_confidence)
}



```



```{r}

alpha <- .1
lambda <- 10
components <- 10L

#factor mtrices for train and test

X_train_conf <- lin_conf(X_train, alpha)
X_test_history_conf <- lin_conf(X_test_history, alpha)



#dim(X_test_history_conf)
```


```{r}
# Initialize a model
# Define hyper parameters
#rank=the number of latent factors in the model (defaults to 10)
#value of alpha is calculated using cross validation





model <- WRMF$new(rank = components,
                  lambda = lambda,
                  feedback = 'implicit',
                  solver = 'conjugate_gradient')




# #testing
# head(X_test_history)
# head(X_test_history_conf,100)
# head(X_train_conf)
# dim(X_train_conf)
# dim(X_test_history_conf)
# class(X_test_history_conf)
# class(X_test_history_conf)
```



```{r}


# Calculate user factors
train_user_factors <- model$fit_transform(X_train_conf)

# Products matrix and recommendations are  made by selecting the top 10  items for which P(ui) is greatest for an user.

test_predictions <- model$predict(X_test_history_conf, k = 10)



new_user_embeddings <- model$transform(X_test_history_conf)
new_user_1 <-  X_test_history_conf[1:1, , drop = FALSE]

new_user_predictions <- model$predict(new_user_1, k = 10)



#Loss and Score or fixed product factors
trace = attr(train_user_factors, "trace")
ggplot(trace) +
  geom_line(aes(x = iter, y = value, col = scorer)) +
  labs(title = "Loss and Scoring Metrics by iteration") +
  theme(plot.title = element_text(hjust = .5))


```


```{r}

mcname<-colnames(X_test_history_conf)
length(mcname)




users <- data.frame(user_id = c("13726"))
prod<-data.frame(product_id =mcname)
ord<-data.frame(user_id = c("13726"),product_id=c("196","130","13176"))


ord$RowIdx <- match(ord$user_id, users$user_id)
ord$ColIdx <- match(ord$product_id, prod$product_id)




matSparse3 <- sparseMatrix(
  i = ord$RowIdx, 
  j = ord$ColIdx, 
  x = 1L,
  dims = c(nrow(users), nrow(prod)), 
  dimnames = list(users$user_id, prod$product_id)
)




# m3<-cbind(1,104608,mcname,1:7463)
# colnames(m3)<-c("uid","user_id","product_id","pid")
# m3<-as.data.frame(m3)
# 
# 
# 
# m3$orders[m3$product_id==21903]<-1
# m3$orders[m3$product_id==11520]<-1
# m3$orders[m3$product_id==39812]<-1
# m3$orders[m3$product_id==25146]<-1
# 
# 
# #m3[is.na(m3)] <- '.'
# 
# 
# m3$uid<-as.integer(m3$uid)
# m3$user_id<-as.integer(m3$user_id)
# m3$product_id<-as.integer(m3$product_id)
# m3$pid<-as.integer(m3$pid)
# m3$orders<-as.integer(m3$orders)







# 
# user_enc<-m3 %>% 
#           distinct(user_id)
# product_enc<-m3 %>% 
#   distinct(product_id)





# m4<-cbind(1,mcname)
# colnames(m4)<-c("user_id","product_id")
# m4<-as.data.frame(m4)
# m4$orders[m4$product_id==21903]<-1
# 
# m4$user_id<-as.integer(m4$user_id)
# m4$product_id<-as.integer(m4$product_id)
# m4[is.na(m4)] <- 0
# dim(m4)

# m5 <- with(m3, sparseMatrix(i=uid, 
#                        j=pid, 
#                        x=orders,
#                        dimnames=list(levels(user_id), levels(product_id))))
# m5


# m7<-cbind(mcname)
# colnames(m7)<-c("product_id")
# m7<-as.data.frame(m7)
# #m7$user_id<-as.integer(m7$user_id)
# m7$product_id<-as.integer(m7$product_id)
# m7$orders[m7$product_id==21903]<-1


# 
# anu<-m7%>%pivot_wider(names_from = product_id,values_from=orders)
# class(anu)
# anupama<-data.matrix(anu)
# class(anupama)
# B <- as(anupama, "sparseMatrix")


# 
# y= sparseMatrix(i = m3$uid, j =m3$pid, x = m3$orders,
#                 dimnames = list(user_enc$user_id, product_enc$product_id))
# 
# X_test_history_conf
# 
# 
# y_shiny <- lin_conf(y, alpha)
# new_user_predictions <- model$predict(y_shiny, k = 10)



new_user_embeddings <- model$transform(X_test_history_conf)
new_user_1 <-  X_test_history_conf[1:1, , drop = FALSE]

new_user_predictions <- model$predict(matSparse3, k =3)

top_list = as(new_user_predictions, "list")

# class(top_list)
# 
# recomdf=data.frame(top_list)
# colnames(recomdf)="product_id"


table_recom<-products %>% 
             filter(product_id %in% top_list) %>% 
            select(product_id,product_name)
print()


class(new_user_predictions)
dim(new_user_predictions)

products %>% filter(product_id %in% c(196,130,13176))

```

```{r}
also_buy<-products %>% filter(product_id %in% c(388,103,112,197,26))
also_buy
```



